---
title: "A3: SPL Library Checkouts"
output: html_document
---

### Introduction

This data has been pulled from [Seattle Open Data](https://data.seattle.gov/Community/Checkouts-by-Title/tmmm-ytt6). I chose to observe the changes in checkout materials over the course of the last 5-6 years, the most popular publishers, the average checkouts each title has, and the publication year that has accumulated the most checkouts. For the checkout material trends, I was very curious to see how the numbers changed before, during, and after COVID. I was interested in learning which publishers are the most popular based on the checkouts because I realized I didn't know many other publishers beside Penguin. The average checkouts and publication year with the most checkouts was something I didn't have a hypothesis going into so I wanted to learn a bit more about this aspect.

### Summary Information

The 5 values that I collected were :
Average number of checkouts: 23.47
Month with the most checkouts: 6
Year with the most checkouts: 2020
Publisher with the most checkouts: Random House, Inc.
Digital checkouts vs. physical checkouts in 2020: ~1.8 million vs. 500k

### The Dataset

The data was collected and provided by the Seattle Public Library, and published by Seattle Open Data. The parameters of the data are number of checkouts, publishers, month, and year. The data was collected from the records of checkouts from the locations of the Seattle Public Library which were compiled together to create this dataset. The data was collected to bring more data to the public. Privacy could be a possible consideration in this data set, though the dataset does not include any personal data to trace it back to a person so it is completely anonymous. I didn't run into many limitations while working with this data, though I did try to find the top five subjects and saw that most books had multiple subjects in the entry, which made finding the most popular a bit more difficult, so I ended up switching to finding the most popular publishers based on the number of checkouts as it was simpler. The publishers were not a list like the different entries in the subject column. Besides the issue with the subjects column, I didn't really run into much else that affected my ability to successfully come up with values that made sense.

### First Trends Over Time Chart

```{r echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("ggplot2")
library("plotly")
spl_data <- read.csv("/Users/zoesteers/Downloads/2017-2023-10-Checkouts-SPL-Data.csv")

ebook_checkouts_per_year <- spl_data %>% 
  filter(MaterialType == "EBOOK") %>% 
  group_by(CheckoutYear, MaterialType) %>% 
  summarise(checkouts_per_year = sum(Checkouts))

physical_book_checkouts_per_year <- spl_data %>% 
  filter(MaterialType == "BOOK") %>% 
  group_by(CheckoutYear, MaterialType) %>% 
  summarise(checkouts_per_year = sum(Checkouts))

ggplot() +
  geom_line(data = physical_book_checkouts_per_year, 
            aes(x = CheckoutYear, 
                y = checkouts_per_year, 
                color = MaterialType)) +
  geom_line(data = ebook_checkouts_per_year, 
            aes(x = CheckoutYear, 
                y = checkouts_per_year, 
                color = MaterialType)) +
  geom_point(data = ebook_checkouts_per_year, 
             aes(x = CheckoutYear, 
                 y = checkouts_per_year)) +
  geom_point(data = physical_book_checkouts_per_year, 
             aes(x = CheckoutYear, 
                 y = checkouts_per_year)) +
  scale_x_continuous(breaks =seq(2017, 2023, 1)) +
  labs(title = "Physical Book and Ebook Trends",
       x = "Checkout Year",
       y = "Number of Checkouts",
       )
```

This chart illustrates the changes in physical book vs ebook checkouts from 2017-2023. We can see the very obvious change beginning in 2020 which coincides with the start of the COVID-19 pandemic. The sharp drop in physical book checkouts and the spike in ebook checkouts can be attributed to the need for social distancing and public spaces, like the library, not being open. Ebook checkouts can be made remotely, while physical book checkouts have to be done in person, which further explains the change that can be observed.

### Second Trends Over Time Chart

```{r echo=FALSE, message=FALSE, warning=FALSE}
library("dplyr")
library("ggplot2")
library("plotly")
library("scales")

digital_usage_class <- spl_data %>% 
  filter(UsageClass == "Digital") %>% 
  group_by(UsageClass, CheckoutYear) %>%
  summarize(digital_checkouts_per_year = sum(Checkouts))

physical_usage_class <- spl_data %>% 
  filter(UsageClass == "Physical") %>% 
  group_by(UsageClass, CheckoutYear) %>%
  summarize(physical_checkouts_per_year = sum(Checkouts))

ggplot() +
  geom_line(data = physical_usage_class,
            aes(x = CheckoutYear,
                y = physical_checkouts_per_year,
                color = UsageClass)) +
  geom_line(data = digital_usage_class,
            aes(x = CheckoutYear,
                y = digital_checkouts_per_year,
                color = UsageClass)) +
  scale_x_continuous(breaks =seq(2017, 2023, 1)) +
  labs(title = "Usage Class Trends",
       x = "Checkout Year",
       y = "Number of Checkouts",
  ) +
  scale_y_continuous(labels = label_number_si())
```

This graph is similar to the last one, but instead takes into account ALL digital and physical checkouts. This includes movies, music, magazines, video disks, audiobooks, and more. This shows how COVID-19 affected even more than only the checkouts of physical books and ebooks, but also many more materials. The number of checkouts overall is much higher, and while the trends look quite similar, there are changes that are more dramatic as a result of more materials being counted.

### Your Choice

```{r echo=FALSE, message=FALSE, warning=FALSE}
popular_publishers <- spl_data %>% 
  group_by(Publisher) %>% 
  summarize(top_checkouts = sum(Checkouts, na.rm = TRUE)) %>%
  arrange(desc(top_checkouts)) %>% 
  filter(!Publisher %in% c(" ", "NULL")) %>% 
  filter(row_number() <= 5)
bar_plot <- ggplot(data = popular_publishers) +
  geom_col(mapping = aes(
    x = Publisher,
    y = top_checkouts,
    fill = Publisher,
    text = paste("Checkouts:", top_checkouts))
    ) + 
  labs(title = "Top Publishers",
             x = "Publishers",
             y = "Number of Checkouts") +
  scale_y_continuous(labels = label_number_si())
ggplotly(bar_plot, tooltip = "text")
```

This bar chart shows the publishers of books that were checked out the most overall. The popularity of certain publishing companies is illustrated in the figure.